---
title: "MTH6139 Time Series"
subtitle: "Coursework 1"
author:
- name: Adesola Dada
date: "March 2024"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r, echo=FALSE}
htmltools::img(src = knitr::image_uri("QMlogo.png"), 
               alt = 'logo', 
               style = 'position:absolute; top:0; right:0; padding:10px; width:30%;')
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Meta’s Prophet forecasting system

I will use Meta’s Prophet forecasting system to predict future values of data. I will be using the "co2" time series which contains 486 monthly observations of the Mauna Loa atmospheric concentrations of CO2.

```{r}
#install.packages("prophet")
# This installs the forecasting system that will be used.


co2.dataframe= data.frame(
  ds=zoo::as.yearmon(time(co2)),
  y=co2) #This makes the "co2" time series a data frame, which is needed to use the forecasting system.

head(co2.dataframe) 
```

Here are the first few data points in the data frame.

```{r message=FALSE, warning=FALSE}
model= prophet::prophet(co2.dataframe) #This runs prophet with the data frame.

future=prophet::make_future_dataframe(model, periods=12, freq="quarter") #This creates future dates for forecasting (precisely this predicts the future 12 quarters).

prediction= predict(model, future) #This uses the forecasting model to predict future values.

plot(model, prediction) 
```

This shows the forecasted values in blue on a plot.

```{r}
prophet::prophet_plot_components(model, prediction)
```

As we can see, there is a strongly positive correlation within the data, with clear yearly seasonality.

Now I will forecast with the same 3 year prediction, but in years instead of quarters.
```{r message=FALSE, warning=FALSE}
future2=prophet::make_future_dataframe(model, periods=3, freq="year") #This creates future dates for forecasting (precisely this predicts the future 3 years).

prediction2= predict(model, future2) #This uses the forecasting model to predict future values.

plot(model, prediction2) 
```

This clearly shows that although we predicted the same period, due to the seasonality within the year, a smaller time period will produce more accurate/ believable results.
